---
title: "Clustering and Differential Expression, RNA-seq, SU2C adult GBM lines"
output:
  html_document:
    toc: true
    df_print: paged
---

## Setup Directories + Load Data

```{r, results='hide'}
## get required packages
library(SummExpDR)
library(DESeq2)
library(SummarizedExperiment)
## Setup directories
top.dir <- '/home/owenwhitley/projects/su2c_v2/'
scripts.dir <- file.path(top.dir, 'scripts/RNA/analysis')
RNA.results.dir <- file.path(top.dir, 'results/RNA')
analysis.results.dir <- file.path(RNA.results.dir, 'RNA_GSCs_clustering_DE')

overwrite <- T
if (overwrite) {
  if (dir.exists(analysis.results.dir)) {
    system(paste('rm -rf', analysis.results.dir))
  }
}

for (dir.name in c(RNA.results.dir, analysis.results.dir)) {
  if (!dir.exists(dir.name)) {
    dir.create(dir.name)
  }
}
setwd(scripts.dir)
```

```{r}
data.dir <- file.path(top.dir, 'data/preprocessed/RNA/SU2C_coh1234_nonSU2C/')
## Load data
load(file = file.path(data.dir, 'RNA_adult_GBM_vst_combat_nonSU2C.RData'))
load(file = file.path(data.dir, 'RNA_adult_GBM_dds.RData'))
load(file = file.path(data.dir, 'RNA_adult_GBM_fpkm_combat_nonSU2C.RData'))
dir(data.dir)

## run checks
stopifnot(all(rownames(RNA_adult_GBM_dds) == rownames(RNA_adult_GBM_vst_combat_nonSU2C)))
stopifnot(all(rownames(RNA_adult_GBM_fpkm_combat_nonSU2C) == rownames(RNA_adult_GBM_vst_combat_nonSU2C)))
stopifnot(all(colnames(RNA_adult_GBM_dds) == colnames(RNA_adult_GBM_vst_combat_nonSU2C)))
stopifnot(all(colnames(RNA_adult_GBM_fpkm_combat_nonSU2C) == colnames(RNA_adult_GBM_vst_combat_nonSU2C)))
stopifnot(identical(colData(RNA_adult_GBM_fpkm_combat_nonSU2C), colData(RNA_adult_GBM_vst_combat_nonSU2C)))
stopifnot(all(sapply(colnames(colData(RNA_adult_GBM_dds)), FUN = function(x) {
  if (any(is.na(colData(RNA_adult_GBM_dds)[,x]))) {
  # skiep NA entries in DESeq2 dataset object
    y <- T
  } else {
    y <- all(colData(RNA_adult_GBM_dds)[,x] == colData(RNA_adult_GBM_vst_combat_nonSU2C)[,x])
  }
})))

summ.expt <- SummarizedExperiment(assays = list(counts = counts(RNA_adult_GBM_dds),
                                                vst = assay(RNA_adult_GBM_vst_combat_nonSU2C, 1),
                                                log2.fpkm = log2(assay(RNA_adult_GBM_fpkm_combat_nonSU2C, 1))),
                                  colData = colData(RNA_adult_GBM_vst_combat_nonSU2C))

## subset for GSC lines
summ.expt <- summ.expt[,which(summ.expt$Type_bis == 'L')]
summ.expt

## add nonSU2C status as batch variable

RNA.GSC.obj <- SummExpDR::create_SummExpDR(summ.expt)

```

## Run PCA + clustering

``` {r}

# remove rows with no variance. If you don't do this, run_pca will throw and error from call to prcomp
RNA.GSC.obj <- SummExpDR::detect_outliers(RNA.GSC.obj, assay_use = 'vst')

RNA.GSC.obj <- SummExpDR::runPCA(RNA.GSC.obj, i = 'vst', std_norm = TRUE)

SummExpDR::plotDR(RNA.GSC.obj, 'PC1', 'PC2', 'PCA', color_by = 'Cohort', shape_by = 'Lab')
SummExpDR::plotDR(RNA.GSC.obj, 'PC1', 'PC2', 'PCA', color_by = 'is.outlier', shape_by = 'Lab')

SummExpDR::screePlot(RNA.GSC.obj, key = 'PCA', dims_use = paste0('PC', 1:20))

RNA.GSC.obj <- SummExpDR::run_multi_k_functions(x = RNA.GSC.obj, analysis_name = 'spectral', 
                                               k_use = 2:4, num_iter = 200, thresh_ARI = 0.80,
                                               thresh_proportion = 0.80, assay_use = 'vst',
                                               algorithm = 'spectral_kernlab', std_norm = T,
                                               sigma = 0.5, affi_K = 10)
RNA.GSC.obj <- SummExpDR::get_soln_k(RNA.GSC.obj, add_2_metadata = T, k = 2, pick_best_valid = F)

SummExpDR::plot_sim_dist(RNA.GSC.obj, save_file = F, mode = 'full.vs.rs', clust_analysis = 'spectral')
SummExpDR::plot_sim_dist(RNA.GSC.obj, save_file = F, mode = 'resample.vs.perm', clust_analysis = 'spectral')

SummExpDR::plotDR(RNA.GSC.obj, 'PC1', 'PC2', 'PCA', color_by = 'spectral_k2', shape_by = 'Lab')

```


We see that k = 2 is the best solution for the data (robust to resampling of data),
so we'll perform differential expression + GSEA between the two things

## Differential Expression + GSEA

#### Run Differential Expression
``` {r}

# run differential expression
RNA.GSC.obj <- SummExpDR::run_diff_exp(expt_data = RNA.GSC.obj, class_type_use = 'spectral_k2', covariates = c('is.nonSU2C'),
                                        class1 = 1, class2 = 2, pipeline = 'DESEQ2', write_output = T,
                                        output_dir = analysis.results.dir, prefix = 'RNA_GSCs_')
analysis_name <- 'RNA_GSCs_spectral_k2_1_v_2'
diff.exp.res <- getAnalyses(RNA.GSC.obj, key = analysis_name)$diff_exp
is.upreg <- diff.exp.res$log2FoldChange > 0
is.downreg <- diff.exp.res$log2FoldChange < 0
passes.fdr <- diff.exp.res$padj < 0.05
c1.upreg <- rownames(diff.exp.res[is.upreg & passes.fdr, ])
c2.upreg <- rownames(diff.exp.res[is.downreg & passes.fdr, ])

# geneset.dir <- file.path(top.dir, '/data/raw/GeneSets/owen_su2c_RNA_oct_2018/')
# c1.markers.old <- read.csv(file.path(geneset.dir, 'RNA_SU2C_coh1234_nonSU2C_GSC_combat_nonSU2C_upreg_markers_k2_clust1.csv'),
#                            stringsAsFactors = F)
# c2.markers.old <- read.csv(file.path(geneset.dir, 'RNA_SU2C_coh1234_nonSU2C_GSC_combat_nonSU2C_upreg_markers_k2_clust2.csv'),
#                            stringsAsFactors = F)
# head(c1.markers.old)
# head(c2.markers.old)
# c1.markers.old <- c1.markers.old$x
# c2.markers.old <- c2.markers.old$x
# 
# length(c1.upreg)
# length(c1.markers.old)
# length(intersect(c1.upreg, c1.markers.old))
# 
# length(c2.upreg)
# length(c2.markers.old)
# length(intersect(c2.upreg, c2.markers.old))

# save 'marker genesets' in results directory
write.csv(c1.upreg, file = file.path(analysis.results.dir, 'spectral_k2_c1_upreg_FDR_0_05.csv'), quote = F, row.names = F)
write.csv(c2.upreg, file = file.path(analysis.results.dir, 'spectral_k2_c2_upreg_FDR_0_05.csv'), quote = F, row.names = F)

```


#### Run ssgsea using markersets
``` {r}
# Will have to add utility for adding metadata
gset.list <- list(RNA.c1.upreg = c1.upreg,
                  RNA.c2.upreg = c2.upreg)
RNA.GSC.obj <- SummExpDR::run_gsva(RNA.GSC.obj, assay = 'log2.fpkm',
                                  genesets = gset.list, method = 'ssgsea',
                                  parallel.sz = 20, verbose = F)

# save(RNA.GSC.ssgsea.result, file = file.path(analysis.results.dir, 'RNA_GSC_ssgsea_result.RData'))
SummExpDR::plotDR(RNA.GSC.obj, 'PC1', 'PC2', 'PCA', color_by = 'RNA.c1.upreg', shape_by = 'spectral_k2')
SummExpDR::plotDR(RNA.GSC.obj, 'PC1', 'PC2', 'PCA', color_by = 'RNA.c2.upreg', shape_by = 'spectral_k2')
SummExpDR::plotDR(RNA.GSC.obj, 'PC1', 'PC2', 'PCA', color_by = 'RNA.c1.upreg', shape_by = 'Lab')
SummExpDR::plotDR(RNA.GSC.obj, 'PC1', 'PC2', 'PCA', color_by = 'RNA.c2.upreg', shape_by = 'Lab')
```

#### Run GSEA
``` {r}
dir(analysis.results.dir)

files <- dir(file.path(analysis.results.dir, 'RNA_GSCs_spectral_k2_1_v_2'))
rankfile <- files[grep('.rnk', files)]
gmt_file <- '../../../data/raw/gmt_files/baderlab_gmts/Human_GO_AllPathways_no_GO_iea_April_01_2018_symbol.gmt'
SummExpDR::enrich_GSEA(gsea_rank_list_path = file.path(analysis.results.dir, 'RNA_GSCs_spectral_k2_1_v_2',
                                            rankfile),
                      gsea_analysis_name = 'RNA_GSCs_c1_v_c2',
                      gsea_out = analysis.results.dir,
                      gsea_gmx = gmt_file)

```

#### Examine results 

``` {r}
dir.contents <- dir(analysis.results.dir)
GSEA.path <- file.path(analysis.results.dir, dir.contents[grep('GseaPreranked', dir.contents, ignore.case = T)])
if (length(GSEA.path) != 1) {
  stop(paste('expected 1 match for GSEA.path'))
}

GSEA.results <- retrieve_gsea(GSEA.path)
```

__Examine Enriched Pathways__
``` {r}
## Positively Enriched
GSEA.results$na_pos_report
## Negatively Enriched
GSEA.results$na_neg_report
```

#### Save Results
__Save Clustering + DiffExp Results to File__
``` {r}
save(RNA.GSC.obj, file = file.path(analysis.results.dir, 'RNA_GSC_obj.RData'))
```

## Session Info

``` {r}
Sys.time()
sessionInfo()
```
