---
title: "Run Logistic Regression to Distinguish Btw Tumour + Line"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 5
params:
  inp_data: '/home/owenwhitley/projects/su2c_v2/data/preprocessed/scRNA/GSCs_Tumour_combined/laura_BTSC_Tumour_combined_Oct_2019_no_G800_L.rds'
  results_dir: '/home/owenwhitley/projects/su2c_v2/results/scRNA/laura_SU2C_GSCs_GBM_combined_Oct_2019_no_G800_L_logistic'
  rerun_logistic: TRUE
  n_splits: 30
  split_fraction: 0.20
  base_seed: 32
  coef_thresh: 0.10
  file_prefix: 'logistic_regression_T_v_L'
  heatmap_fontsize: 10.0
  num_dims: 2
  other_reductions_plot: 'pca'
  dims_use_mapping: 'NULL'
  reduction_use: 'pca'
  feats_plot: 'NULL'
---
# Overview

Here we are interested in seeing if tumour and line cells are linearly separable in PCA space and will
be training a logistic regression classifier to see if this is the case. We will examine which PCs
the classifier considers important, and see what genes are associated with said PCs.

# Setup

``` {r}
suppressPackageStartupMessages({
  library(ComplexHeatmap)
  source('./whitley_scRNA_helpers.R')
})
for (i in 1:length(params)) {
  param.i <- names(params)[i]
  assign(param.i, params[[i]])
  if (param.i == 'inp_data') {
    if (is.character(inp_data)) {
      # we only load data in from file for testing/debugging purposes.
      # inp_data should generally be a Seurat object that is fully preprocessed 
      # as outlined in 'seurat_pipeline' function in helper functions,
      # with PCA calculated
      if (file.exists(inp_data)) {
        inp_data <- readRDS(inp_data)
      } else {
        stop('file does not exist')
      }
    }
  }
}
# setup results directory if does not exist
if (!dir.exists(results_dir)) {
  dir.create(results_dir)
}

other_reductions_plot <- setdiff(other_reductions_plot, reduction_use)

# not current run options
run_opts_curr <- params[c('n_splits', 'split_fraction', 'base_seed', 'file_prefix')]
results_file <- paste0(file_prefix, '.pydata')
run_opts_file <- paste0(file_prefix, '_run_opts.rds')

# setup mapping of dim reduction names to dims to use for each dim reduction in 'other_reductions
if (dims_use_mapping == 'NULL') {
  # dims_use_mapping should be a named list of lists, with each 'sublist' containing pairs of 
  # dimensions to plot
  dims_use_mapping <- list(pca = list(c(1,2), c(1,3)),
                           tsne = list(c(1,2)))
} else {
  if (!is.list(dims_use_mapping)) {
    stop('dims_use_mapping, if specified, must be a list')
  }
}
if (!all(other_reductions_plot %in% names(dims_use_mapping))) {
  stop('all reductions plotted must be in names of dim_use_mapping')
}
# decide whether or not we will be plotting feature loadings for pca. may change in future to accomodate mnn/fastmnn loadings
use_pca <- 'pca' == reduction_use

# specify features to plot if left unspecified
if (feats_plot == 'NULL') {
  feats_plot <- c('SampleType', 'pred_class', 'misclassified', 'SampleID')
}

```

# Computations

```{r}

if (!file.exists(file.path(results_dir, results_file)) || rerun_logistic) {
  # see if there are any PCs that can separate tumour + line
  T_v_L_logistic <- fit_logistic_multi(X = GetCellEmbeddings(inp_data, 
                                                             reduction.type = reduction_use)[,1:num_dims],
                                       y = as.character(inp_data@meta.data$SampleType),
                                       num_splits = n_splits, 
                                       split_fraction = split_fraction, 
                                       base_seed = base_seed,
                                       penalty = 'l2')
  # save run options
  run_opts_prev <- run_opts_curr
  reticulate::py_save_object(T_v_L_logistic, file = file.path(results_dir, results_file))
  saveRDS(run_opts_prev, file = file.path(results_dir, run_opts_file))
} else {
  T_v_L_logistic <- reticulate::py_load_object(file.path(results_dir, results_file))
  run_opts_prev <- readRDS(file.path(results_dir, run_opts_file))
  if (!identical(run_opts_prev, run_opts_curr)) {
    warning(run_opts_prev)
    stop('previous run options not identical to current run options. \
         reset run options to previous values or set rerun_logistic to TRUE')
  }
}

gc(full = TRUE)
  
```

# View Results

## Train  + Test Accuracy, Model Coefficients

``` {r, fig.width = 5, fig.height = 7}

# plot train + test Accuracy
train_test_accuracy <- data.frame(Data = factor(c(rep('Train', n_splits), rep('Test', n_splits)), levels = c('Train', 'Test')),
                                  Accuracy = c(T_v_L_logistic$train_accuracy_vals,
                                               T_v_L_logistic$test_accuracy_vals))

p <- ggplot(data = train_test_accuracy, mapping = aes(Data, Accuracy)) + geom_boxplot()
p <- p + 
  ggtitle('Train + Test Accuracy\n30 splits') + 
  theme(text = element_text(size = 20), plot.title = element_text(hjust = 0.5)) +
  theme_cowplot()
print(p)

# plot model coefficients
split_coefs <- lapply(1:ncol(T_v_L_logistic$coefs_mat), 
                      FUN = function(x)return(T_v_L_logistic$coefs_mat[,x]))
coefs_df <- data.frame(dimension = numeric(0), coef = numeric(0))
for (i in 1: length(split_coefs)) {
  new_df <- data.frame(dimension = rep(i, length(split_coefs[[i]])),
                       coef = split_coefs[[i]])
  coefs_df <- rbind(coefs_df, new_df)
  rm(new_df)
}
coefs_df$dimension <- factor(coefs_df$dimension)
xlab_model_coef <- paste(toupper(reduction_use), 'Dimension')
p <- ggplot(data = coefs_df, mapping = aes(dimension, coef)) + geom_boxplot()
p <- p + 
  ggtitle('Model Coefficients\n30 splits') + 
  theme(text = element_text(size = 20), plot.title = element_text(hjust = 0.5)) +
  xlab(xlab_model_coef) + ylab('Coefficient') + 
  theme_cowplot()
print(p)

model_coefs <- T_v_L_logistic$final_model$model$coef_
model_coefs
```

## Look at Model Coefficients

``` {r}
# filter PCs by having model coefficients > thresh
interesting_dims <- which(abs(model_coefs) > coef_thresh)
```

``` {r, fig.height = 12, fig.width = 6, dev = 'svg', eval=use_pca}


gene_loadings <- Seurat::GetDimReduction(inp_data, 
                                         reduction.type = 'pca', 
                                         slot = 'gene.loadings')
head(gene_loadings)
# only look at gene loadings for genes in top 20 for at least 1 PC in PCs of interest
if (length(interesting_dims) >= 1) {
  gene_loadings <- gene_loadings[,interesting_dims]
  return_top_20 <- function(x) {
    x[order(abs(x), decreasing = TRUE)][1:20]
  }
  
  if (class(gene_loadings) == 'matrix') {
    in_top_genes <- apply(gene_loadings, MARGIN = 2, FUN = function(x) {
    names(x) <- rownames(gene_loadings)
    return(names(x) %in% names(return_top_20(x)))
    })
    genes_keep <- apply(in_top_genes, FUN = function(x) {any(x)}, MARGIN = 1)
    gene_loadings <- gene_loadings[genes_keep,]
  } else if (is.vector(gene_loadings)) {
    in_top_genes <- names(gene_loadings) %in% names(return_top_20(gene_loadings))
    gene_loadings <- gene_loadings[in_top_genes]
    gene_loadings <- matrix(gene_loadings, ncol = 1, 
                            dimnames = list(rownames = names(gene_loadings), colnames = interesting_dims[1]))
  }
  
  Heatmap(gene_loadings, row_names_gp = gpar(fontsize = heatmap_fontsize), name = 'PC loading',
          cluster_columns = FALSE)
  
  # PCHeatmap(inp_data, pc.use = interesting_dims)

} else {
  print('No PCs observed with coefficients above threshold')
}

```

## Visualize Where Predicted Classes Occur on Dim Red Plots

### Make Predictions and Label Misclassified Cells

``` {r}
cell_embeddings <- Seurat::GetCellEmbeddings(inp_data, reduction = reduction_use)[,1:num_dims]
pred_class <- T_v_L_logistic$final_model$model$predict(cell_embeddings)
inp_data@meta.data$pred_class <- pred_class
inp_data@meta.data$misclassified <- inp_data@meta.data$pred_class != inp_data@meta.data$SampleType
saveRDS(inp_data, file = file.path(results_dir, paste0(file_prefix, '_seurat.rds')))
```

### Imput Embedding Space Plots

``` {r}
try({
  if (length(interesting_dims) > 1) {
    for (i in 1:(length(interesting_dims) - 1)) {
      for (j in (i+1):length(interesting_dims)) {
        make_dr_plots(inp_data, dim1 = interesting_dims[i], dim2 = interesting_dims[j],
                      feats_plot = feats_plot, reduction.use = reduction_use, legend_pt_size = 5)
        cat('**Misclassified Cells Only**\n')
        make_dr_plots(inp_data, dim1 = interesting_dims[i], dim2 = interesting_dims[j],
                      feats_plot = feats_plot, reduction.use = reduction_use,
                      filter_by = 'misclassified', filter_class = 'TRUE',
                      legend_pt_size = 5)
      }
    }
  } else {
    make_dr_plots(inp_data, dim1 = 1, dim2 = 2,
                  feats_plot = feats_plot, reduction.use = reduction_use,
                  legend_pt_size = 5)
    cat('**Misclassified Cells Only**\n')
    make_dr_plots(inp_data, dim1 = 1, dim2 = 2,
                  feats_plot = feats_plot, reduction.use = reduction_use,
                  filter_by = 'misclassified', filter_class = 'TRUE',
                  legend_pt_size = 5)
  }
})



```

### other reductions to use

``` {r, results='asis'}
for (dr_name in other_reductions_plot) {
  dr_name_upper <- toupper(dr_name)
  cat(paste('####', dr_name_upper, '\n'))
  dims_use_list <- dims_use_mapping[[dr_name]]
  
  for (i in 1:length(dims_use_list)) {
    dims_i <- dims_use_list[[i]]
    cat(paste('#####', paste(dr_name_upper, dims_i[1]), paste(dr_name_upper, dims_i[2]), '\n'))
    try({
      make_dr_plots(inp_data, dim1 = dims_i[1], dim2 = dims_i[2], 
                      feats_plot = feats_plot, reduction.use = dr_name, legend_pt_size = 5)
      cat('**Misclassified Cells Only**\n')
      make_dr_plots(inp_data, dim1 = dims_i[1], dim2 = dims_i[2], 
                        feats_plot = feats_plot, reduction.use = dr_name,
                    filter_by = 'misclassified', filter_class = 'TRUE', legend_pt_size = 5)
    })
    cat('\n')
    cat('\n')
  }
}

```


## Assess Proportions of Tumour + Line Cells in Misclassified Cells

``` {r, error=TRUE}

stacked_barcharts(inp_data = inp_data@meta.data, split_by = 'misclassified', color_by = 'SampleType')
stacked_barcharts(inp_data = inp_data@meta.data, split_by = 'misclassified', color_by = 'SampleType', show_proportion = TRUE)
stacked_barcharts(inp_data = inp_data@meta.data, split_by = 'SampleType', color_by = 'misclassified')
stacked_barcharts(inp_data = inp_data@meta.data, split_by = 'SampleType', color_by = 'misclassified', show_proportion = TRUE)

# make table displaying misclassified cells by SampleType (Tumour or Line)
sampletype_vs_misclassified <- make_table(inp_data@meta.data, var1 = 'SampleType', var2 = 'misclassified')
sampletype_vs_misclassified
write.csv(sampletype_vs_misclassified, file = file.path(results_dir, paste0(file_prefix, '_sampletype_vs_misclassified.csv')))

# show misclassified cells by Sample ID
# tumours
p <- stacked_barcharts(inp_data = inp_data@meta.data[inp_data@meta.data$SampleType == 'Tumour',], 
                       split_by = 'SampleID', color_by = 'misclassified', show_proportion = FALSE)
p <- p + theme(axis.text.x = element_text(angle = 90, size = heatmap_fontsize))
print(p)
p <- stacked_barcharts(inp_data = inp_data@meta.data[inp_data@meta.data$SampleType == 'Tumour',], 
                       split_by = 'SampleID', color_by = 'misclassified', show_proportion = TRUE)
p <- p + theme(axis.text.x = element_text(angle = 90, size = heatmap_fontsize))
print(p)

# lines
p <- stacked_barcharts(inp_data = inp_data@meta.data[inp_data@meta.data$SampleType == 'Line',], 
                       split_by = 'SampleID', color_by = 'misclassified', show_proportion = FALSE)
p <- p + theme(axis.text.x = element_text(angle = 90, size = heatmap_fontsize))
print(p)
p <- stacked_barcharts(inp_data = inp_data@meta.data[inp_data@meta.data$SampleType == 'Line',], 
                       split_by = 'SampleID', color_by = 'misclassified', show_proportion = TRUE)
p <- p + theme(axis.text.x = element_text(angle = 90, size = heatmap_fontsize))
print(p)
# make tables showing same info
SampleID_vs_misclassified <- make_table(inp_data@meta.data, var1 = 'SampleID', var2 = 'misclassified', filter_by = NULL, filter_class = NULL)
# order by tumour/line status
SampleID_vs_misclassified <- SampleID_vs_misclassified[order(grepl('_T', rownames(SampleID_vs_misclassified))),]
SampleID_vs_misclassified
write.csv(SampleID_vs_misclassified, file = file.path(results_dir, paste0(file_prefix, '_SampleID_vs_misclassified.csv')))
write.csv(SampleID_vs_misclassified, file = file.path(results_dir, paste0(file_prefix, '_SampleID_vs_misclassified.csv')))

```

# SessionInfo

``` {r}
Sys.time()
sessionInfo()
```

